---
title: "Using woody to run wood transport models"
author: "Lise Vaudor"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{woody}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", cache=TRUE
)
```

# Prepare data

```{r tree_mypath}
mypath="../../woody_data/wood_data_train/"
mystation="V2942010"
fs::dir_tree(mypath)
```

```{r attach_woody}
library(woody)
```

Import all wood log files inside a file (1 directory per event)

```{r import_Wdata, message=FALSE, warning=FALSE}
Wdat=woody::import_Wdata(mypath) 
``` 

Collect qtvar data from banquehydro, for the period covered by Wdata:

```{r bh_get_qtvar}
Qdat=woody::get_Qdata(station=mystation,
                      Wdata=Wdat,
                      extratime=1200)
``` 



Now complete Qdata with variables T_Q and S

```{r complete_Qdata}
Qdatc=woody::complete_Qdata(Qdata=Qdat)
```

Complete Wdat with discharge data (interpolating Qdat)

```{r complete_Wdata_with_Qdata}
Wdatc=woody::complete_Wdata_with_Qdata(Wdata=Wdat,
                                       Qdata=Qdatc)
```

Then calculate Wdata_wt, where 1 row= 1 waiting time between two wood occurrences.

```{r Wdata_as_waiting_times}
Wdat_wt=woody::Wdata_as_waiting_times(Wdatc)

```

# Run random forest

```{r}
myrf <- run_rf(Wdat_wt)
myrf
```

Predict on training data itself:

```{r}
Wdat_wt=woody::predict_rf(obj_rf=myrf,
                          newdata=Wdat_wt)
```

R2:

```{r}
calc_rf_R2(Wdat_wt)
```

# Get predictions on another dataset

```{r, warning=FALSE, message=FALSE}
pathpred="../../woody_data_predict/"
mystation="V2942010"
Wdat_pred=import_Wdata(pathpred)
Qdat_pred=get_Qdata(station=mystation,
                       Wdata=Wdat_pred)
Qdat_pred=complete_Qdata(Qdat_pred)
Wdat_pred=complete_Wdata_with_Qdata(Wdat_pred,Qdat_pred)
Wdat_pred=Wdata_as_waiting_times(Wdat_pred)
Wdat_pred=predict_rf(myrf,Wdat_pred)
calc_rf_R2(Wdat_pred)
```

