---
title: "woody: predict wood occurrence data based on discharge data"
author: "Lise Vaudor"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{woody}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Prepare data

```{r tree_mypath}
mypath="../data-raw/mem_data/woody_data_train/"
mystation="V2942010"
fs::dir_tree(mypath)
```

```{r attach_woody}
library(woody)
```

Import all wood log files inside a file (1 directory per event)

```{r import_Wdata, message=FALSE, warning=FALSE}
Wdat=import_Wdata(mypath) 
``` 

Collect **qtvar data** from banquehydro, for the **period covered by Wdata**

```{r bh_get_qtvar}
Qdat=get_Qdata(Wdata=Wdat,
               station=mystation,
               back_in_time=TRUE)

Qdat %>% 
  dplyr::pull(Time) %>% 
  range()

head(Qdat)
```

Now **complete Qdata** with **variables T_Q and S**

```{r complete_Qdata}
Qdatc=complete_Qdata(qtvar=Qdat,site="site1")
head(Qdatc)
```

Complete Wdat with discharge data (interpolating Qdat)

```{r complete_Wdata_with_Qdata}
Wdatc=complete_Wdata_with_Qdata(Wdata=Wdat,
                                Qdata=Qdatc)
head(Wdatc)
```

Then calculate Wdata_wt, where 1 row= 1 waiting time between two wood occurrences.

```{r Wdata_as_waiting_times}
Wdat_wt=Wdata_as_waiting_times(Wdatc)
```

# Run random forest

```{r run_rf}
myrf <- run_rf(Wdat_wt)
myrf
```

Predict on training data itself

```{r pred_on_train}
Wdat_wt=predict_rf(obj_rf=myrf,
                   newdata=Wdat_wt)
```

R2:

```{r calc_R2_on_train}
calc_rf_R2(Wdat_wt)
```

# Get predictions on another dataset

```{r pred_on_other_data, warning=FALSE, message=FALSE}
pathpred="../data-raw/mem_data/woody_data_predict/"

Wdat_forpred=import_Wdata(pathpred)

Qdat_forpred= get_Qdata(Wdat_forpred, station="V2942010") %>% 
    complete_Qdata()

Wdat_pred=Wdat_forpred %>%
  complete_Wdata_with_Qdata(Qdat_forpred) %>% 
  Wdata_as_waiting_times() 

pred_rf=predict_rf(Wdat_pred,myrf)
calc_rf_R2(pred_rf)
```

