---
title: "1. woody: calculate flux"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  rmarkdown::html_vignette:
    toc: true           
    number_sections: true   
    df_print: kable
vignette: >
  %\VignetteIndexEntry{1. woody: calculate flux}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r attach_woody, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(warning=FALSE,message=FALSE)
library(woody)
library(tidyverse)
library(DiagrammeR)
```

# Global process

## For one site

```{r diagramme0, echo=FALSE, fig.height=10}
grViz("
digraph boxes_and_circles {
  graph [overlap = true, fontsize = 10]
  node [shape = box, style= filled]
  A [fillcolor=pink]
  Qc  [fillcolor = cadetblue1]
  W [fillcolor = burlywood1]
  Qh [fillcolor=cadetblue1]
  myrf[fillcolor=gold]
  Qp [fillcolor=cadetblue2]
  Qn [fillcolor=cadetblue3]
  Wn [fillcolor=burlywood2]
  N  [fillcolor=darkkhaki]
  
  node [shape = ellipse,fixedsize = false,color = grey, style= filled, fillcolor=grey90]
  'interp_Qdata()'
  'predict_rf()'
  'summarise_Qdata()'  
  'summarise_Wdata()'
  
  # several 'edge' statements
  Qc -> 'interp_Qdata()' [arrowhead = none]
  'interp_Qdata()'-> Qh [label= 'interpolate Qdata hourly']
  Qh, myrf-> 'predict_rf()'
  'predict_rf()' -> 'Qp' [label = ' add Y_pred values']
  Qp -> 'summarise_Qdata()' [arrowhead = none]
  'summarise_Qdata()' -> 'Qn' [label = ' add N_pred values']
  A,W -> 'summarise_Wdata()' [arrowhead = none]
  'summarise_Wdata()' -> 'Wn' [label = ' add N_obse values']
  Wn,Qn  -> 'N' 

}
")
```

## For several sites

Here, we want to carry out this process for two sites, so that we use and enrich a nested table from step to step in the following manner:

```{r diagramme1, echo=FALSE, fig.height=7}
grViz("
digraph boxes_and_circles {
  graph [overlap = true,  labeljust=l]
  
  graph [fontsize=10 compound=true overlap=true];
    node [shape=egg fontsize=10, style=filled]
    // all nodes related to Q
    node [fillcolor=cadetblue1]
    'Qc8'  [label='Qc[1,2]'] Qh
    'Qh' [label='Qh[1,2]']
    node [fillcolor=cadetblue2]
    'Qp'  [label='Qp[1,2]'] 
    node [fillcolor=cadetblue3]
    'Qn'  [label='Qn[1,2]']
    'Qn12'  [label='Qn[1,2]']
    'Qn13'  [label='Qn[1,2]']
    // all nodes related to W
    node [fillcolor=burlywood1, label='W[1,2]']
    'W8' 'W9' 'W10' 'W11'
    // all nodes related to W+Qnode 
     node [label= 'W[1,2]', fillcolor = burlywood]
    'W8' 'W9' 'W10' 'W11'
    node [label= 'Wn[1,2]', fillcolor = burlywood2]
    'Wn' 'Wn13'
    node [label= 'N[1,2]', fillcolor = darkkhaki]
    'N'
    
    node[shape=box]
    myrf [fillcolor = gold, label='myrf']
    
    // all nodes related to descriptors
    node [shape=egg, fillcolor=ivory, label='descriptors[1,2]']
    'descriptors8'    'descriptors9'    'descriptors10' 'descriptors11' 'descriptors12' 'descriptors13'
    
    subgraph cluster_8 {'descriptors8' 'W8' 'Qc8' label='tib_WpQc'}
    subgraph cluster_9 {'descriptors9' 'W9' 'Qh'; label='tib_WpQh'}
    subgraph cluster_10 {'descriptors10' 'W10' 'Qp' ; label='tib_WpQp'}
    subgraph cluster_11 {'descriptors11' 'W11' 'Qn'; label='tib_WpQn'}
    subgraph cluster_12 {'descriptors12' 'Wn' 'Qn12'; label='tib_WnQn'}
    subgraph cluster_13 {'descriptors13' 'Wn13', 'Qn13' 'N'; label='tib_WnQnN'}
    
    edge[color=grey]
    'descriptors8' -> 'descriptors9' -> 'descriptors10' -> 'descriptors11' -> 'descriptors12' -> 'descriptors13'
    'W8' -> 'W9' -> 'W10' -> 'W11' -> 'Wn' -> 'Wn13'
    'descriptors11' -> 'Wn' [label='An[1,2]', fontsize=6]
    'Qc8' -> 'Qh' 
    'Qp' -> 'Qn' -> 'Qn12' -> 'Qn13'  
    'Qh','myrf' -> 'Qp'
    'Wn','Qn12' -> 'N'
  }
")
```

# From log-flux estimates $Y_{pred}$ to flux estimates $N_{pred}$

First, let's see how the **predictions of log-flux** $Y_{pred}$ (estimated through the Random Forest) might be used to calculate **predictions of flux $N_{pred}$**. It is indeed necessary to **take into account the residuals error of the model** to do so.

$$Y =log(N)= log(3600/W)$$

where :

- $N$ is an hourly flux of wood
- $W$ corresponds to the waiting time(in seconds) between two wood occurrences
- 3600 corresponds to 3600 seconds i.e. one hour

## Distribution of $Y_{pred}$

Let's have a look at the **distribution of residuals**:

```{r plot_distrib, fig.width=6}
tib_WpQc=readRDS("../data-raw/results/tib_WpQc.RDS")
Wdata_pred=tib_WpQc %>% 
  select(Wdata) %>% 
  tidyr::unnest(Wdata, .name_repair="minimal")
p=ggplot(Wdata_pred,aes(x=Y_pred,y=Y-Y_pred))+
  geom_point(alpha=0.05)+
  facet_grid(cols=vars(site),
             labeller = labeller(.rows = label_both,
                                 .cols = label_both))+
  geom_hline(aes(yintercept=0), col='red')
plot(p)
```

The observed value $Y$ is distributed around predicted value $Y_{pred}$, in an **approximately Gaussian distribution with standard deviation $\sigma$**, hence we can hypothesize that

$$Y \sim \mathcal{N}(\mu,\sigma) $$

$$E(Y)=E(Y_{pred})=\mu$$

$\sigma$, as the **residual standard deviation**, corresponds to a measure of error of the model predicting $Y=log(N)$.

Here is an estimate of $\sigma$ for both sites:

```{r calc_sigma}
Wdata_pred %>% 
  group_by(site) %>% 
  mutate(residuals=Y_pred-Y) %>% 
  summarise(sigma=sd(residuals))
```


For now, we will estimate roughly that $\sigma=1.5$ for both sites.

On the log scale, $E(Y)=E(Y_{pred})$, but going back to a **natural (not logged) scale for fluxes**:

$$E(N)=E(e^Y)=e^{E(Y)+\frac{\sigma^2}{2}}$$

Hence $E(N)>e^{E(Y)}$: the expected value of $N$, $E(N)$,  is higher than the exponential of $E(Y)$.

## Empirical explanation of $E(N_{pred})>exp(E(Y_{pred}))$

This can also be understood, in a **more empirical way**, like this:

The error in $log(N)$ is symmetrical, which means that $N$ can equally be **a bit lower** or **a lot higher** than predicted (on a natural scale). Hence the mean of values on this natural scale is higher that just the exponential of the mean on the log scale.

In the following section, we will hence use this **estimate of $\sigma$ to convert $Y_{pred}$ estimates to $N_{pred}$ estimates**.

# Summarise `Qdata` hourly and calculate $N_{pred}$ and $N_{obs}$

## Interpolate `Qdata` hourly

Let's **interpolate Qdata at an hourly timescale** and use these interpolated, hourly data to calculate $Y_{pred}$.

```{r tib_WQh}
myrf=readRDS("../data-raw/results/rf.RDS")
tib_WQc=readRDS("../data-raw/results/tib_WQc.RDS")
tib_WQh=tib_WQc %>%
  # interpolate at an hourly timescale
  mutate(Qdata=purrr::map(.x=Qdata,
                          ~interp_Qdata(.x, timescale="hours")))
```

## Calculate hourly $Y_pred$

```{r tib_WQp}
tib_WQp=tib_WQh %>%
  mutate(Qdata=purrr::map(.x=Qdata %>% na.omit(),
                          ~predict_rf(newdata=.x,obj_rf=myrf)))

```

## Deduce hourly $N_{pred}$

```{r tib_WQn}
tib_WQn=tib_WQp %>%
  mutate(apath=paste0("../data-raw/annot_times_",site,".csv")) %>%
  mutate(Qdata=purrr::map(.x=Qdata,
                          ~summarise_Qdata(Qdata=.x,sigma=1.5)))
```

## Get hourly $N_{obs}$ and $N_{obse}$ from `Wdata`

We want to compare these values $N_{pred}$ with the **observed number of wood pieces during an hour** $N_{obse}$.

The **observation of wood pieces is not continuous**, so that we have to take into account, for each hour, the **proportion of time the observation of wood pieces really occurred** $P_{obs}$.  This information is retrievable from `Adata` (A for "annotation").

$N_{obse}=N_{obs}/P_{obs}$



```{r tib_WnQn}
tib_WnQn=tib_WQn %>%
  mutate(Adata=purrr::map2(.x=apath,.y=site,
                           ~import_Adata(path=.x,site=.y))) %>%
  mutate(Wdata=purrr::map2(.x=Wdata,.y=Adata,
                           ~summarise_Wdata(Wdata=.x,Adata=.y)))
```

Once imported and cleaned, this how `Adata` (first lines, for the Ain site) looks like:

```{r show_Adata, echo=FALSE}
knitr::kable(head(tib_WnQn$Adata[[1]]))
```

So this is how `Wdata` now looks like, with a new **extrapolated estimate for the observed flux** $N_{obse}$ (first lines, for the Ain site)


```{r show_tib_WnQn_Wdata, echo=FALSE}
knitr::kable(head(tib_WnQn$Wdata[[1]]))
```

## Join values $N_{pred}$ (from `Qdata`) and $N_{obse}$ (from `Wdata`) into a single table `Ndata`


```{r tib_WnQnN}
tib_WnQnN=tib_WnQn %>% 
  mutate(Ndata=purrr::map2(.x=Qdata,.y=Wdata,
                           ~left_join(.x,.y,by=c("site","Time"))))
```

# Summary 

To make the production of plots easier, we unnest `Ndata` from `tib_WnQnN`:

```{r unnest_Ndata}
Ndata=tib_WnQnN %>% 
  ungroup() %>% 
  select(Ndata) %>% 
  tidyr::unnest(Ndata)
```

## Model performance

```{r R2}
R2=tib_WnQnN %>% select(site,Ndata) %>% 
  mutate(R2=purrr::map_df(Ndata,calc_rf_R2,type="N")) %>% 
  tidyr::unnest(R2)
R2
```



## Plot of $N_{pred}$ and $N_{obs}$


```{r plot_N, fig.height=8, fig.width=8}
Ndataplot=Ndata %>% 
  tidyr::pivot_longer(cols=c(N_pred,N_obse), names_to="type") %>%
  filter(!is.na(event))
ggplot(Ndataplot, aes(x=Time,y=value))+
  geom_path(aes(col=type))+
  geom_point(aes(col=type))+
  facet_wrap(site~event, scales="free", ncol=2)
```
