---
title: "woody"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  rmarkdown::html_vignette:
    toc: true           
    number_sections: true   
    df_print: kable
vignette: >
  %\VignetteIndexEntry{woody}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(woody)
library(tidyverse)
library(DiagrammeR)
```

# Global process

```{r diagramme, echo=FALSE, fig.height=10}
grViz("
digraph boxes_and_circles {
  graph [overlap = true, fontsize = 10]
  node [shape = box, style= filled]
  wpath [fillcolor = burlywood1]
  qpath  [fillcolor = cadetblue1]
  W [fillcolor = burlywood1]
  Q  [fillcolor = cadetblue1]
  Wc [fillcolor = darkkhaki]
  Qc  [fillcolor = cadetblue1]
  Q [fillcolor = cadetblue1]
  Wwt [fillcolor = darkkhaki]
  Wwt_test [fillcolor = darkkhaki]
  myrf [fillcolor = gold]
  Wp [fillcolor = darkolivegreen2]
  
  node [shape = ellipse,fixedsize = false,color = grey, style= filled, fillcolor=grey90]
  'import_Wdata()'
  'import_Qdata()'
  'complete_Qdata()' 
  'complete_Wdata_with_Qdata()'
  'Wdata_as_waiting_times()'
  'run_rf()'
  'predict_rf()'

  # several 'edge' statements
  wpath -> 'import_Wdata()' [arrowhead = none]
  'import_Wdata()'-> W 
  qpath -> 'import_Qdata()' [arrowhead = none]
  'import_Qdata()'-> Q 
  Q -> 'complete_Qdata()' [arrowhead=none]
  'complete_Qdata()' -> Qc 
  W -> 'complete_Wdata_with_Qdata()' [arrowhead=none]
  Qc-> 'complete_Wdata_with_Qdata()' [arrowhead=none]
  'complete_Wdata_with_Qdata()' -> Wc
  Wc -> 'Wdata_as_waiting_times()' [arrowhead=none]
  'Wdata_as_waiting_times()' -> Wwt
  Wwt -> 'run_rf()' [arrowhead = none]
  'run_rf()' -> myrf 
  myrf -> 'predict_rf()' [arrowhead = none]
  'Wwt_test' -> 'predict_rf()' [arrowhead = none]
  'predict_rf()' -> 'Wp' 
}
")
```

# Prepare data

## Get `W` data about wood occurrences

```{r diagramme1, echo=FALSE, fig.height=3}
grViz("
digraph boxes_and_circles {
  graph [overlap = true, fontsize = 10]
  node [shape = box, style= filled]
  wpath [fillcolor = burlywood]
  W [fillcolor = burlywood]

  node [shape = ellipse,fixedsize = false,color = grey, style= filled, fillcolor=grey90]
  'import_Wdata()'
 

  # several 'edge' statements
  wpath -> 'import_Wdata()' [arrowhead = none]
  'import_Wdata()'-> W 
}
")
```

```{r tree_mypath}
wpath="../data-raw/mem_data/woody_data_train/"
mystation="V2942010"
fs::dir_tree(wpath)
```

```{r attach_woody}
library(woody)
```

Import all wood log files inside a file (1 directory per event)

```{r import_Wdata, message=FALSE, warning=FALSE}
W=import_Wdata(wpath,site="training") 
``` 

## Get `Q` data about discharge

```{r diagramme2, echo=FALSE, fig.height=3}
grViz("
digraph boxes_and_circles {
  graph [overlap = true, fontsize = 10]
  graph [overlap = true, fontsize = 10]
  node [shape = box, style= filled]
  qpath  [fillcolor = cadetblue1]
  Q  [fillcolor = cadetblue1]

  
  node [shape = ellipse,fixedsize = false,color = grey, style= filled, fillcolor=grey90]
  'import_Qdata()'
  
  qpath -> 'import_Qdata()' [arrowhead = none]
  'import_Qdata()'-> Q 

}
")
```

Import `Q`:

```{r bh_get_qtvar}
qpath="../data-raw/Qdata/Qdatc_Ain.csv"
Q=import_Qdata(path=qpath, site="Ain")
```

The first lines of `Q` look like this

```{r show_Q, echo=FALSE}
knitr::kable(head(Q))
```

## Complete `Q` into `Qc`

```{r diagramme3, echo=FALSE, fig.height=3}
grViz("
digraph boxes_and_circles {
   graph [overlap = true, fontsize = 10]
  node [shape = box, style= filled]
  Qc  [fillcolor = cadetblue1]
  Q [fillcolor = cadetblue1]

  
  node [shape = ellipse,fixedsize = false,color = grey, style= filled, fillcolor=grey90]
  'complete_Qdata()' 


  # several 'edge' statements
  Q -> 'complete_Qdata()' [arrowhead=none]
  'complete_Qdata()' -> Qc 

}
")
```

Now **complete `Q`** with **variables T_Q and S**

```{r complete_Qdata}
result_file="../data-raw/results/Q.RDS"
if(!file.exists(result_file)){
  Qc=complete_Qdata(qtvar=Q,site="Ain")
  saveRDS(Qc,result_file)
}
Qc=readRDS(result_file)
```
The first lines of `Qc` look like this:

```{r show_Qc, echo=FALSE}
knitr::kable(head(Qc))
```

## Complete `W` with `Q` and get `Wc`

```{r diagramme4, echo=FALSE, fig.height=3}
grViz("
digraph boxes_and_circles {
 graph [overlap = true, fontsize = 10]
  node [shape = box, style= filled]
  W [fillcolor = burlywood]
  Wc [fillcolor = darkkhaki]
  Qc  [fillcolor = cadetblue1]
  
  node [shape = ellipse,fixedsize = false,color = grey, style= filled, fillcolor=grey90]
  'complete_Wdata_with_Qdata()'


  # several 'edge' statements
  W -> 'complete_Wdata_with_Qdata()' [arrowhead=none]
  Qc-> 'complete_Wdata_with_Qdata()' [arrowhead=none]
  'complete_Wdata_with_Qdata()' -> Wc
}
")
```

Complete `W` with discharge data (interpolating `Q`)

```{r complete_Wdata_with_Qdata}
Wc=complete_Wdata_with_Qdata(Wdata=W,
                             Qdata=Qc)
```

The first lines of `Wc` look like this:

```{r show_Wc, echo=FALSE}
knitr::kable(head(Wc))
```

## Transform `Wc` into `Wwt`

```{r diagramme5, echo=FALSE, fig.height=3}
grViz("
digraph boxes_and_circles {
  graph [overlap = true, fontsize = 10]
  node [shape = box, style= filled]
  Wc [fillcolor = darkkhaki]
  Wwt [fillcolor = darkkhaki]
 
  node [shape = ellipse,fixedsize = false,color = grey, style= filled, fillcolor=grey90]
  'Wdata_as_waiting_times()'

  # several 'edge' statements
 
  Wc -> 'Wdata_as_waiting_times()' [arrowhead=none]
  'Wdata_as_waiting_times()' -> Wwt
}
")
```

Then calculate `Wwt`, where 1 row= 1 waiting time between two wood occurrences.

```{r Wdata_as_waiting_times}
Wwt=Wdata_as_waiting_times(Wc)
```

The first lines of `Wwt` look like this:

```{r show_Wwt, echo=FALSE}
knitr::kable(head(Wwt))
```

# Run random forest

```{r diagramme6, echo=FALSE, fig.height=3}
grViz("
digraph boxes_and_circles {
   graph [overlap = true, fontsize = 10]
  node [shape = box, style= filled]
  Wwt [fillcolor = darkkhaki]
  myrf [fillcolor = gold]
  
  node [shape = ellipse,fixedsize = false,color = grey, style= filled, fillcolor=grey90]

  'run_rf()'

  # several 'edge' statements
  Wwt -> 'run_rf()' [arrowhead = none]
  'run_rf()' -> myrf 
}
")
```

```{r run_rf}
myrf <- run_rf(Wwt)
myrf
```

# Get predictions and assess random forest performance

```{r diagramme7, echo=FALSE, fig.height=3}
grViz("
digraph boxes_and_circles {
   graph [overlap = true, fontsize = 10]
  node [shape = box, style= filled]
  Wwt_test [fillcolor = darkkhaki]
  myrf [fillcolor = gold]
  Wp [fillcolor = darkolivegreen2]
  
  node [shape = ellipse,fixedsize = false,color = grey, style= filled, fillcolor=grey90]
  'predict_rf()'

  # several 'edge' statements
 
  Wwt_test -> 'predict_rf()' [arrowhead = none] 
  myrf -> 'predict_rf()' [arrowhead = none]
  'predict_rf()' -> 'Wp' 
}
")
```

## On the training dataset itself


Predict on training data itself

```{r pred_on_train}
Wp=predict_rf(obj_rf=myrf,
              newdata=Wwt)
```

R2:

```{r calc_R2_on_train}
calc_rf_R2(Wp)
```

## On another dataset

```{r pred_on_other_data, warning=FALSE, message=FALSE}
wpathpred="../data-raw/mem_data/woody_data_predict/"

Wwt_test=import_Wdata(wpathpred) %>%
  complete_Wdata_with_Qdata(Qc) %>% 
  Wdata_as_waiting_times() 

Wp_test=predict_rf(Wwt_test,myrf)
calc_rf_R2(Wp_test)
```

