---
title: "Using woody to run wood transport models"
author: "Lise Vaudor"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{woody}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", cache=TRUE
)
```

# Prepare data

```{r tree_mypath}
mypath="../../woody_data/woody_data_train/"
mystation="V2942010"
fs::dir_tree(mypath)
```

```{r attach_woody}
library(woody)
```

Import all wood log files inside a file (1 directory per event)

```{r import_Wdata, message=FALSE, warning=FALSE}
Wdat=import_Wdata(mypath) 
``` 

Collect qtvar data from banquehydro, for the period covered by Wdata

```{r bh_get_qtvar}
Qdat=get_Qdata(Wdata=Wdat,
               station=mystation,
               extratime=1200)
``` 



Now complete Qdata with variables T_Q and S

```{r complete_Qdata}
Qdatc=complete_Qdata(Qdata=Qdat)
```

Complete Wdat with discharge data (interpolating Qdat)

```{r complete_Wdata_with_Qdata}
Wdatc=complete_Wdata_with_Qdata(Wdata=Wdat,
                                Qdata=Qdatc)
```

Then calculate Wdata_wt, where 1 row= 1 waiting time between two wood occurrences.

```{r Wdata_as_waiting_times}
Wdat_wt=Wdata_as_waiting_times(Wdatc)
```

# Run random forest

```{r run_rf}
myrf <- run_rf(Wdat_wt)
myrf
```

Predict on training data itself

```{r pred_on_train}
Wdat_wt=predict_rf(obj_rf=myrf,
                   newdata=Wdat_wt)
```

R2:

```{r calc_R2_on_train}
calc_rf_R2(Wdat_wt)
```

# Get predictions on another dataset

```{r, warning=FALSE, message=FALSE}
pathpred="../../woody_data/woody_data_predict/"
mystation="V2942010"

Wdat_forpred=pathpred %>% 
    import_Wdata()

Qdat_forpred=Wdat_forpred %>%
    get_Qdata(station=mystation) %>% 
    complete_Qdata()

Wdat_pred=Wdat_forpred %>%
  complete_Wdata_with_Qdata(Qdat_forpred) %>% 
  Wdata_as_waiting_times(Wdat_forpred) %>% 
  predict_rf(myrf)

Wdat_pred %>%
  calc_rf_R2()
```

